### 2023-03-13 main.tex
duality

The process described by the generator \eqref{Generator} is reversiblewith respect to the homogeneus product measure \begin{equation}\label{reversibleMeasure}
\mu_{rev}=\bigotimes_{x\in V}\mu^{rev}_{x}

@@

\mu_{rev}^{x}(n^{x})=\frac{\nu}{\prod_{k=1}^{N}n_{k}^{x}}\prod_{k=1}^{N}p_{k}^{n_{k}^{x}}

@@

\noindent
\textbf{Remark}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\mu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process characterized by the generator \eqref{Generator}. The state space \eqref{stateSpace} is the vector space with basis elements

@@

\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}\label{edgeHamiltonian}
H_{x,y}=\sum_{k,l=1}^{N}\left(E_{kl}^{x}\otimes E_{lk}^{y}-E_{ll}^{x}\otimes E_{kk}^{y}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{k,l=1}^{N}\alpha_{k}^{x}\left(E_{kl}^{x}-E_{ll}^{x}\right)
\end{equation}
Here $E_{kl}^{x}$ is a copy of $E_{kl}$ defined in \eqref{actionE} acting on site $x$. 
Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by the following relation:
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

@@

We can write this Hamiltonian as a function of the coproduct of the second Casimir
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
that acts diagonally as $C_{2}|\mathbf{n}\rangle=\twoj(\twoj+N)|\mathbf{n}\rangle$. \\
 Considering two sites $x,y\in \mathcal{E}$ the standard coproduct operator is defined as 
\begin{equation}
\begin{split}
\Delta^{x,y}:gl(N)&\to gl(N)\otimes gl(N)\\
E_{ab}&\to E_{ab}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes E_{ab}^{y}
\end{split}
\end{equation}
therefore, we have 
\begin{equation}
\Delta^{x,y}(C_{2})=\sum_{a,b=1}^{N}\Delta(E_{ab})\Delta(E_{ba})=2\sum_{a,b=1}^{N}E_{ab}^{x}\otimes E_{ba}^{y}+C_{2}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes C_{2}^{y}
\end{equation}

@@

\section{Duality}
\subsection{Definition of duality}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$ if
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
\textbf{Remark}:
in the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the processes are self-dual with respect to a self-duality function.
\newline
\newline
\textbf{Remark}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining transformation between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads as
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
where $\widetilde{H}=\widetilde{\mathcal{L}}^{T}$.
\subsection{Duality for the generalized multi-species boundary driven stirring process}
We now state our duality result. The process $|n\rangle$ with Hamiltonian \eqref{} is dual in the sense of \eqref{DualityRelation}, with respect to a process $|\xi\rangle$ defined on the enlarged state space
\begin{equation}
    \widetilde{\Omega}:=\Omega\times \mathbb{N}_{0}^{V}
\end{equation}
with elements
\begin{equation}
    |\xi\rangle=\bigotimes_{x\in V}\left(|r_{1}^{x},\ldots,r_{N}^{x}\rangle\otimes |m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}\rangle\right)
\end{equation}
where $r_{x}^{k}$ denotes the number of dual particles of type $k\in \{1,\ldots,N\}$ at site $x\in V$ and $m_{k}^{\widetilde{x}}$ denotes the number of particles of species $k\in \{2,\ldots,N\}$ at an extra-site $\widetilde{x}$ connected to the site $x\in V$.\\
The Hamiltonian of the dual process reads as
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}

>>>>

The process described by the generator \eqref{Generator} is reversible with respect to the homogeneus product measure \begin{equation}\label{reversibleMeasure}
\mu_{rev}=\bigotimes_{x\in V}\mu_{rev}^{x}

@@

\mu_{rev}^{x}(n^{x})=\frac{\nu!}{\prod_{k=1}^{N}n_{k}^{x}!}\prod_{k=1}^{N}p_{k}^{n_{k}^{x}}

@@

\noindent
\textbf{Remark}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\mu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process with generator \eqref{Generator}. The state space \eqref{stateSpace} is the vector space with basis elements

@@

\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}\label{edgeHamiltonian}
H_{x,y}=\sum_{k,l=1}^{N}\left(E_{kl}^{x}\otimes E_{lk}^{y}-E_{ll}^{x}\otimes E_{kk}^{y}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{k,l=1}^{N}\alpha_{k}^{x}\left(E_{kl}^{x}-E_{ll}^{x}\right)
\end{equation}
Here $E_{kl}^{x}$ is a copy of $E_{kl}$ defined in \eqref{actionE} acting on site $x$. 
Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

@@

We can write the edge Hamiltonian \eqref{edgeHamiltonian} as a function of the coproduct of the second Casimir of $gl(N)$
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
that acts diagonally as $C_{2}|\mathbf{n}\rangle=\twoj(\twoj+N)|\mathbf{n}\rangle$ and belogs to the center of $gl(N)$ (i.e. it commutes with all the basis elements).  \\
 Considering two sites $x,y\in \mathcal{E}$ the standard coproduct operator is defined as 
\begin{equation}
\begin{split}
\Delta^{x,y}:gl(N)&\to gl(N)\otimes gl(N)\\
E_{ab}&\to E_{ab}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes E_{ab}^{y}
\end{split}
\end{equation}
therefore, we have 
\begin{equation}
\Delta^{x,y}(C_{2})=\sum_{a,b=1}^{N}\Delta(E_{ab})\Delta(E_{ba})=2\sum_{a,b=1}^{N}E_{ab}^{x}\otimes E_{ba}^{y}+C_{2}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes C_{2}^{y}
\end{equation}

@@

\section{Duality}
\subsection{Definition of duality}
Consider two Markov processes: $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$ if
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
\textbf{Remark}:
in the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the processes are self-dual with respect to a self-duality function.
\newline
\newline
\textbf{Remark}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining transformation between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads as
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
where $\widetilde{H}=\widetilde{\mathcal{L}}^{T}$.
\subsection{Duality for the generalized multi-species boundary driven stirring process}
We now state our duality result. The process $|n\rangle$ with Hamiltonian \eqref{} is dual in the sense of \eqref{DualityRelation}, with respect to a process $|\xi\rangle$ defined on the enlarged state space
\begin{equation}
    \widetilde{\Omega}:=\Omega\times \mathbb{N}_{0}^{V}
\end{equation}
with elements
\begin{equation}
    |\xi\rangle=\bigotimes_{x\in V}\left(|r_{1}^{x},\ldots,r_{N}^{x}\rangle\otimes |m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}\rangle\right)
\end{equation}
where $r_{x}^{k}$ denotes the number of dual particles of type $k\in \{1,\ldots,N\}$ at site $x\in V$ and $m_{k}^{\widetilde{x}}$ denotes the number of particles of species $k\in \{2,\ldots,N\}$ at an extra-site $\widetilde{x}$ connected to the site $x\in V$.\\
The Hamiltonian of the dual process reads as
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
